---
title: "R Notebook"
output: html_notebook
---
This is our dataset before pre-processing.
```{r}
library(readxl)
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")
print(CreditCard)
```
```{r}
# Load the necessary libraries (if not already loaded)
library(dplyr)
library(readxl)
library(ggplot2)  # Load the ggplot2 library for plotting

# Load your own dataset from an Excel file (replace 'your_dataset.xlsx' with your Excel file)
your_data <-  read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Calculate class distribution
class_distribution <- your_data %>%
  group_by(Approved) %>%
  summarise(Count = n())

# Create a bar plot to visualize class distribution
ggplot(class_distribution, aes(x = factor(Approved), y = Count)) +
  geom_bar(stat = "identity") +
  labs(title = "Class Distribution", x = "Class", y = "Count")

```

Data cleaning: our data set already cleaned, there are no incomplete all colomns are filled , inconsistent each colomn have the same type  or intentional data. it has only noisy data (outliers) and we removed them.






Scatter plot shows the relationship between the age and the yearsEmployed if the age increased the years also will increase.
```{r}
#Scatter plot
with(CreditCard, plot(Age,YearsEmployed))
```

Hisogram shows the most frequent debt which is -1.
```{r}
#histogram 
var(CreditCard$Debt)
hist(CreditCard$Debt)
```

Barplot shows how many applicants are citizen or not.
```{r}
#barplot
library(dplyr)
CreditCard$Citizen %>% table() %>% barplot()
```
*Data intgration(correlation analysis) first we will compare the nominal colomns with the class label.

Data intgration(correlation analysis) between the Industry colomn (nominal data type) and the class label colomn approved. 
We calculated the chi square = 98.323 and the df= 13 so we found that the crtical value for df 13 is = 34.528 ,the chi square is larger than the critical value we reject H0 so they are depandent and correlated .
```{r}
contingency_table <- table(CreditCard$Industry, CreditCard$Approved)
chi_square_result <- chisq.test(contingency_table)
print(chi_square_result)
``` 
Data intgration(correlation analysis) between the Ethnicity colomn (nominal data type) and the class label colomn approved. 
We calculated the chi square = 41.813 and the df= 4 so we found that the crtical value for df 4 is = 18.465 ,the chi square is larger than the critical value we reject H0 so they are depandent and correlated .
```{r}
contingency_table <- table(CreditCard$Ethnicity, CreditCard$Approved)
chi_square_result <- chisq.test(contingency_table)
print(chi_square_result)
```
Data intgration(correlation analysis) between the Citizen colomn (nominal data type) and the class label colomn approved. 
We calculated the chi square = 9.1916 and the df= 2 so we found that the crtical value for df 2 is = 13.815 ,the chi square is larger than the critical value we reject H0 so they are depandent and correlated .

```{r}
contingency_table <- table(CreditCard$Citizen, CreditCard$Approved)
chi_square_result <- chisq.test(contingency_table)
print(chi_square_result)
```
after we did chi-square all the columns was highly correlated to we didn't apply feature selection.






*Data intgration(correlation analysis) Second we will compare the numeric colomns with the class label,we found that all numeric coulmns are positive corrleated except zip code coulmn so we need to deleate it.
```{r}
correlation<- cor(CreditCard$Age, CreditCard$Approved)
print(correlation)
```

```{r}
correlation<- cor(CreditCard$Debt, CreditCard$Approved)
print(correlation)
```

```{r}
correlation<- cor(CreditCard$YearsEmployed, CreditCard$Approved)
print(correlation)
```

```{r}
correlation<- cor(CreditCard$CreditScore, CreditCard$Approved)
print(correlation)
```
```{r}
correlation<- cor(CreditCard$Income, CreditCard$Approved)
print(correlation)
```


```{r}
correlation<- cor(CreditCard$ZipCode, CreditCard$Approved)
print(correlation)
```
```{r}
CreditCard<-subset(CreditCard,select= -ZipCode)
print(CreditCard)
```



Data cleaning (noisy outliers)
We relize in Debt coulmn has wide range so it might have outliers
First we detect and found the outliers from the debt colomn .
```{r}
#Find outliers

library(outliers)

OutDebt = outlier(CreditCard$Debt, logical =TRUE)
sum(OutDebt)
Find_outlier = which(OutDebt ==TRUE, arr.ind = TRUE)
OutDebt
Find_outlier
```
Second we deleted the outliers, and then printed the data set after the changes.
```{r}
#remove outliers

CreditCard= CreditCard[-Find_outlier,]
print(CreditCard)
```
Data cleaning (noisy outliers)
As same as Debt ,Income has also huge range.
Here also we detect and found the outliers from the income colomn. .
```{r}
#Find outliers for Income

library(outliers)

OutIncome = outlier(CreditCard$Income, logical =TRUE)
sum(OutIncome)
Find_outlier = which(OutIncome ==TRUE, arr.ind = TRUE)
OutIncome
Find_outlier
```
Second we deleted the outliers from the income colomn, and then printed the data set after the changes.

```{r}
#remove outliers

CreditCard= CreditCard[-Find_outlier,]
print(CreditCard)
```
Data transformation (normalization)
The range of Debt was very wide so we did normalization on Debt to make it in smaller range.
```{r}
#normlization
CreditCard [, 3:3] = scale(CreditCard [, 3:3])
print(CreditCard)

```
Data transformation (encoding)
we noticed that the Income has huge range and The mean = 1017.386 so any value bigger than the mean we encoded it to high and any value smaller than the mean it became low.
```{r}
#Encoding for income
Income_mean<- mean(CreditCard$Income)
print(Income_mean)
CreditCard$Income[CreditCard$Income > Income_mean]<- "High"
CreditCard$Income[CreditCard$Income <= Income_mean]<- "low"
print(CreditCard)
```

This is our Dataset after pre-processing.
```{r}
print(CreditCard)
```
```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (70%) and testing (30%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.7 * nrow(CreditCard))
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using information gain on the training data
tree <- rpart(formula, data = train_data, method = "class", parms = list(split = "information"))

# Visualize the decision tree
rpart.plot(tree)

```





```{r}
# Load required libraries
library(readxl)
library(C50)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (70%) and testing (30%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.7 * nrow(CreditCard))
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Build the decision tree using C5.0 with gain ratio
# Assuming "Approved" is a binary variable (0 or 1)
train_data$Approved <- as.factor(train_data$Approved)
tree <- C5.0(train_data[, -which(names(train_data) == class_attribute)], train_data[[class_attribute]], control = C5.0Control(earlyStopping = FALSE))

# Print the decision tree
plot(tree)

```



```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (70%) and testing (30%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.7 * nrow(CreditCard))
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using Gini index on the training data
tree <- rpart(formula, data = train_data, method = "class", control = rpart.control(cp = 0.005))

# Visualize the decision tree
rpart.plot(tree)


```
```{r}
# Make predictions on the test data
predicted <- predict(tree, test_data, type = "class")

# Calculate confusion matrix
confusion_matrix <- table(Actual = test_data$Approved, Predicted = predicted)

# Calculate accuracy
accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)

# Calculate precision
precision <- confusion_matrix[2, 2] / sum(confusion_matrix[, 2])

# Calculate sensitivity (true positive rate)
sensitivity <- confusion_matrix[2, 2] / sum(confusion_matrix[2, ])

# Calculate specificity (true negative rate)
specificity <- confusion_matrix[1, 1] / sum(confusion_matrix[1, ])

# Print the results
cat("Accuracy:", accuracy, "\n")
cat("Precision:", precision, "\n")
cat("Sensitivity:", sensitivity, "\n")
cat("Specificity:", specificity, "\n")

```



```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (80%) and testing (20%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.8 * nrow(CreditCard))  # Use 0.8 for 80% training
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]


# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using information gain on the training data
tree <- rpart(formula, data = train_data, method = "class", parms = list(split = "information"))

# Visualize the decision tree
rpart.plot(tree)

```

```{r}
# Load required libraries
library(readxl)
library(C50)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (80%) and testing (20%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.8 * nrow(CreditCard))  # Use 0.8 for 80% training
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Build the decision tree using C5.0 with gain ratio
# Assuming "Approved" is a binary variable (0 or 1)
train_data$Approved <- as.factor(train_data$Approved)
tree <- C5.0(train_data[, -which(names(train_data) == class_attribute)], train_data[[class_attribute]], control = C5.0Control(earlyStopping = FALSE))

# Print the decision tree
plot(tree)

```


```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (80%) and testing (20%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.8 * nrow(CreditCard))  # Use 0.8 for 80% training
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using Gini index on the training data
tree <- rpart(formula, data = train_data, method = "class", control = rpart.control(cp = 0.005))

# Visualize the decision tree
rpart.plot(tree)

```
```{r}
# Make predictions on the test data
predicted <- predict(tree, test_data, type = "class")

# Calculate confusion matrix
confusion_matrix <- table(Actual = test_data$Approved, Predicted = predicted)

# Calculate accuracy
accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)

# Calculate precision
precision <- confusion_matrix[2, 2] / sum(confusion_matrix[, 2])

# Calculate sensitivity (true positive rate)
sensitivity <- confusion_matrix[2, 2] / sum(confusion_matrix[2, ])

# Calculate specificity (true negative rate)
specificity <- confusion_matrix[1, 1] / sum(confusion_matrix[1, ])

# Print the results
cat("Accuracy:", accuracy, "\n")
cat("Precision:", precision, "\n")
cat("Sensitivity:", sensitivity, "\n")
cat("Specificity:", specificity, "\n")

```


```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (90%) and testing (10%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.9 * nrow(CreditCard))  # Use 0.9 for 90% training
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using information gain on the training data
tree <- rpart(formula, data = train_data, method = "class", parms = list(split = "information"))

# Visualize the decision tree
rpart.plot(tree)

```


```{r}
# Load required libraries
library(readxl)
library(C50)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (90%) and testing (10%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.9 * nrow(CreditCard))  # Use 0.9 for 90% training
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Build the decision tree using C5.0 with gain ratio
# Assuming "Approved" is a binary variable (0 or 1)
train_data$Approved <- as.factor(train_data$Approved)
tree <- C5.0(train_data[, -which(names(train_data) == class_attribute)], train_data[[class_attribute]], control = C5.0Control(earlyStopping = FALSE))

# Print the decision tree
plot(tree)

```

```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (90%) and testing (10%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.9 * nrow(CreditCard))  # Use 0.9 for 90% training
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using Gini index on the training data
tree <- rpart(formula, data = train_data, method = "class", control = rpart.control(cp = 0.005))

# Visualize the decision tree
rpart.plot(tree)

```
```{r}
# Make predictions on the test data
predicted <- predict(tree, test_data, type = "class")

# Calculate confusion matrix
confusion_matrix <- table(Actual = test_data$Approved, Predicted = predicted)

# Calculate accuracy
accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)

# Calculate precision
precision <- confusion_matrix[2, 2] / sum(confusion_matrix[, 2])

# Calculate sensitivity (true positive rate)
sensitivity <- confusion_matrix[2, 2] / sum(confusion_matrix[2, ])

# Calculate specificity (true negative rate)
specificity <- confusion_matrix[1, 1] / sum(confusion_matrix[1, ])

# Print the results
cat("Accuracy:", accuracy, "\n")
cat("Precision:", precision, "\n")
cat("Sensitivity:", sensitivity, "\n")
cat("Specificity:", specificity, "\n")

```

```{r}
# Load required libraries
library(readxl)
library(cluster)
library(ggplot2)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Select the numeric attributes for clustering (exclude non-numeric or irrelevant columns)
numeric_attributes <- CreditCard[, c("Age", "YearsEmployed", "Debt", "CreditScore", "Income")]

# Specify the range of K values you want to test
k_values <- 1:10  # You can adjust the range

# Initialize a vector to store the WCSS values
wcss_values <- vector()

# Calculate WCSS for each K value
for (k in k_values) {
  kmeans_result <- kmeans(numeric_attributes, centers = k)
  wcss_values <- c(wcss_values, kmeans_result$tot.withinss)
}

# Create a data frame to plot K versus WCSS
elbow_data <- data.frame(K = k_values, WCSS = wcss_values)

# Plot K versus WCSS
ggplot(elbow_data, aes(x = K, y = WCSS)) +
  geom_line() +
  geom_point() +
  labs(x = "Number of Clusters (K)", y = "Within-Cluster Sum of Squares (WCSS)") +
  ggtitle("Elbow Method to Choose K")

```


```{r}
# Load required libraries
library(readxl)
library(dplyr)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the features for clustering
features <- CreditCard[, c("Age", "YearsEmployed", "Debt", "CreditScore", "Income")]

# Specify the values of K (2, 3, and 4)
k_values <- c(2, 3, 4,5)

# Perform K-Means clustering for each value of K
results <- list()

for (k in k_values) {
  set.seed(123)  # Set seed for reproducibility
  kmeans_result <- kmeans(features, centers = k)
  results[[as.character(k)]] <- kmeans_result
}

# Print cluster assignments for each K
for (k in k_values) {
  cat("Clusters for K =", k, ":\n")
  print(results[[as.character(k)]])
  cat("\n")
}


```


```{r}
 #average silhouette for each clusters 
library(cluster)
avg_sil <- silhouette(kmeans.result$cluster,dist(CreditCard)) #a dissimilarity object inheriting from class dist or coercible to one. If not specified, dmatrix must be.
fviz_silhouette(avg_sil)#k-means clustering with estimating k and initializations

```


```{r}
# Specify the numeric features for clustering
library(fpc)
features <- CreditCard[, c("Age", "YearsEmployed", "Debt", "CreditScore", "Income")]

# Perform k-means clustering for a range of K values
k_values <- 2:4  # Define the range of K values to test
silhouette_scores <- numeric(length(k_values))

for (i in 1:length(k_values)) {
  k <- k_values[i]
  kmeans_result <- kmeans(features, centers = k, nstart = 25)  # You can adjust nstart as needed
  
  # Calculate Silhouette score
  silhouette_scores[i] <- silhouette(kmeans_result$cluster, dist(features))
}

# Plot the results
plot(2:4, silhouette_scores, type = "b", xlab = "Number of Clusters (K)", ylab = "Silhouette Score")

# Identify the optimal K
optimal_k <- k_values[which.max(silhouette_scores)]
cat("Optimal number of clusters (K) using Silhouette method:", optimal_k, "\n")

```

