---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
# Credit Card Approval
## Goal :
The goal of collecting of this dataset to predict which people in the dataset are successful in applying for a credit card.

## Data Source :
Here is the data source link It from Kaggle website , The link(
<https://www.kaggle.com/datasets/samuelcortinhas/credit-card-approval-clean-data>) show
the source data.

## the dataset contains 16 attribute and 691 rows.

## Dataset Attributes:
Gender (binary)
Age(numeric interval)
Debt(numeric interval)
Married(binary)
BankCustomer(binary)
industry(nominal)
Ethnicity(nominal)
YearsEmploye

# This is our dataset before pre-processing.

```{r}
library(readxl)
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")
print(CreditCard)
```


## our class label is Approved and as shown in the plot the heights of the bars are roughly equal,so the class label is balanced.

```{r}
# Load the necessary libraries (if not already loaded)
library(dplyr)
library(readxl)
library(ggplot2)  # Load the ggplot2 library for plotting


CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Calculate class distribution
class_distribution <- CreditCard %>%
  group_by(Approved) %>%
  summarise(Count = n())

# Create a bar plot to visualize class distribution
ggplot(class_distribution, aes(x = factor(Approved), y = Count)) +
  geom_bar(stat = "identity") +
  labs(title = "Class Distribution", x = "Class", y = "Count")

```

# Data cleaning:
### our data set already cleaned, there are no incomplete all colomns are filled , inconsistent each colomn have the same type  or intentional data. it has only noisy data (outliers) and we removed them.







### the scatter plot shows the relationship between age and credit score.there are a lot of variation in the data, with some younger people having higher credit scores than older people. This is likely due to a variety of factors, such as income, debt, and spending habits.
```{r}
plot(CreditCard$Age, CreditCard$CreditScore, main = "Scatter Plot: Age vs. CreditScore", xlab = "Age", ylab = "CreditScore", pch = 19)
```

# Hisogram shows the most frequent debt which is -1.
```{r}
#histogram 
var(CreditCard$Debt)
hist(CreditCard$Debt)
```

# Barplot shows how many applicants are citizen or not.
```{r}
#barplot
library(dplyr)
CreditCard$Citizen %>% table() %>% barplot()
```
# Data intgration(correlation analysis) 
### first we will compare the nominal colomns with the class label.

# Data intgration(correlation analysis)
### between the Industry colomn (nominal data type) and the class label colomn approved. 
### We calculated the chi square = 98.323 and the df= 13 so we found that the crtical value for df 13 is = 34.528 ,the chi square is larger than the critical value we reject H0 so they are depandent and correlated .
```{r}
contingency_table <- table(CreditCard$Industry, CreditCard$Approved)
chi_square_result <- chisq.test(contingency_table)
print(chi_square_result)
``` 
# Data intgration(correlation analysis)
### between the Ethnicity colomn (nominal data type) and the class label colomn approved. 
### We calculated the chi square = 41.813 and the df= 4 so we found that the crtical value for df 4 is = 18.465 ,the chi square is larger than the critical value we reject H0 so they are depandent and correlated .
```{r}
contingency_table <- table(CreditCard$Ethnicity, CreditCard$Approved)
chi_square_result <- chisq.test(contingency_table)
print(chi_square_result)
```
# Data intgration(correlation analysis) 
### between the Citizen colomn (nominal data type) and the class label colomn approved. 
### We calculated the chi square = 9.1916 and the df= 2 so we found that the crtical value for df 2 is = 13.815 ,the chi square is larger than the critical value we reject H0 so they are depandent and correlated .

```{r}
contingency_table <- table(CreditCard$Citizen, CreditCard$Approved)
chi_square_result <- chisq.test(contingency_table)
print(chi_square_result)
```
# after we did chi-square all the columns was highly correlated to we didn't apply feature selection.






# Data intgration(correlation analysis) 
### Second we will compare the numeric colomns with the class label,we found that all numeric coulmns are positive corrleated except zip code coulmn so we need to deleate it.
```{r}
correlation<- cor(CreditCard$Age, CreditCard$Approved)
print(correlation)
```

```{r}
correlation<- cor(CreditCard$Debt, CreditCard$Approved)
print(correlation)
```

```{r}
correlation<- cor(CreditCard$YearsEmployed, CreditCard$Approved)
print(correlation)
```

```{r}
correlation<- cor(CreditCard$CreditScore, CreditCard$Approved)
print(correlation)
```
```{r}
correlation<- cor(CreditCard$Income, CreditCard$Approved)
print(correlation)
```


```{r}
correlation<- cor(CreditCard$ZipCode, CreditCard$Approved)
print(correlation)
```

```{r}
CreditCard<-subset(CreditCard,select= -ZipCode)
print(CreditCard)
```



# Data cleaning (noisy outliers)
### We relize in Debt coulmn has wide range so it might have outliers
### First we detect and found the outliers from the debt colomn .
```{r}
#Find outliers

library(outliers)

OutDebt = outlier(CreditCard$Debt, logical =TRUE)
sum(OutDebt)
Find_outlier = which(OutDebt ==TRUE, arr.ind = TRUE)
OutDebt
Find_outlier
```
## Second we deleted the outliers, and then printed the data set after the changes.
```{r}
#remove outliers

CreditCard= CreditCard[-Find_outlier,]
print(CreditCard)
```
# Data cleaning (noisy outliers)
### As same as Debt ,Income has also huge range.
### Here also we detect and found the outliers from the income colomn. .
```{r}
#Find outliers for Income

library(outliers)

OutIncome = outlier(CreditCard$Income, logical =TRUE)
sum(OutIncome)
Find_outlier = which(OutIncome ==TRUE, arr.ind = TRUE)
OutIncome
Find_outlier
```
## Second we deleted the outliers from the income colomn, and then printed the data set after the changes.

```{r}
#remove outliers

CreditCard= CreditCard[-Find_outlier,]
print(CreditCard)
```
# Data transformation (normalization)
### The range of Debt was very wide so we did normalization on Debt to make it in smaller range.
```{r}
#normlization
CreditCard [, 3:3] = scale(CreditCard [, 3:3])
print(CreditCard)

```
# Data transformation (encoding)
### we noticed that the Income has huge range and The mean = 1017.386 so any value bigger than the mean we encoded it to high and any value smaller than the mean it became low.
```{r}
#Encoding for income
Income_mean<- mean(CreditCard$Income)
print(Income_mean)
CreditCard$Income[CreditCard$Income > Income_mean]<- "High"
CreditCard$Income[CreditCard$Income <= Income_mean]<- "low"
print(CreditCard)
```

# This is our Dataset after pre-processing.
```{r}
print(CreditCard)
```

### we split the data into training and testing sets using a 70-30 ratio because it helps preventing overfitting and it provides better real-world performance estimation. 

### then we created a formula for the decision tree using the specified class attribute. Then, we built the decision tree model using the `rpart()` function with the "class" method and information gain as the splitting criterion.

### Finally, we visualized the decision tree using the `rpart.plot()` function.



```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (70%) and testing (30%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.7 * nrow(CreditCard))
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using information gain on the training data
tree_info_gain <- rpart(formula, data = train_data, method = "class", parms = list(split = "information"))

# Visualize the decision tree
rpart.plot(tree_info_gain)

# Build the decision tree using information gain on the training data
tree_info_gain <- rpart(formula, data = train_data, method = "class", parms = list(split = "information"))

# Make predictions on the test data
predicted_info_gain <- predict(tree_info_gain, test_data, type = "class")

# Calculate accuracy
accuracy_info_gain <- sum(predicted_info_gain == test_data$Approved) / nrow(test_data)

# Print the accuracy
cat("Accuracy with Information Gain (70-30 Split)", accuracy_info_gain, "\n")

```

### we again split the data into training and testing sets using a 70-30 ratio. 
### then, we Built the decision tree using C5.0 with gain ratio Assuming that “Approved" is a binary variable (0 or 1)  and printed it.


```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")
# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (70%) and testing (30%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.7 * nrow(CreditCard))
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using gain ratio on the training data
tree_gain_ratio <- rpart(formula, data = train_data, method = "class", parms = list(split = "gain"))

# Visualize the decision tree
rpart.plot(tree_gain_ratio)

# ... (previous code)

# Build the decision tree using gain ratio on the training data
tree_gain_ratio <- rpart(formula, data = train_data, method = "class", parms = list(split = "gain"))

# Make predictions on the test data
predicted_gain_ratio <- predict(tree_gain_ratio, test_data, type = "class")

# Calculate accuracy
accuracy_gain_ratio <- sum(predicted_gain_ratio == test_data$Approved) / nrow(test_data)

# Print the accuracy
cat("Accuracy with Gain Ratio (70-30 Split):", accuracy_gain_ratio, "\n")


```


### we again split the data  using a 70-30 ratio. 
### then created a formula for the decision tree using the specified class attribute. Then, built the decision tree model using the `rpart()` function with the "class" method and Gini index as the splitting criterion.



```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (70%) and testing (30%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.7 * nrow(CreditCard))
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using Gini index on the training data
tree_gini_index <- rpart(formula, data = train_data, method = "class", control = rpart.control(cp = 0.005))

# Visualize the decision tree
rpart.plot(tree_gini_index)


# Build the decision tree using Gini index on the training data
tree_gini_index <- rpart(formula, data = train_data, method = "class", control = rpart.control(cp = 0.005))

# Make predictions on the test data
predicted_gini_index <- predict(tree_gini_index, test_data, type = "class")

# Calculate accuracy
accuracy_gini_index <- sum(predicted_gini_index == test_data$Approved) / nrow(test_data)

# Print the accuracy
cat("Accuracy with Gini Index (70-30 Split):", accuracy_gini_index, "\n")



```
## after doing 3 trees for 70-30 split the best tree with highest accuracy was Gini index




### here we made predictions on the test data
### created the Confusion Matrix
### calculated the Accuracy, Precision, ensitivity (True Positive Rate) , Specificity (True Negative Rate) and printed the results.

### By examining accuracy, precision, sensitivity, and specificity, stakeholders can gain valuable insights into the strengths and weaknesses of the model, aiding in decision-making and potential model refinement.




```{r}
tree <- rpart(Approved ~ ., data = train_data, method = "class")
# Make predictions on the test data
predicted <- predict(tree, test_data, type = "class")

# Calculate confusion matrix
confusion_matrix <- table(Actual = test_data$Approved, Predicted = predicted)

# Calculate accuracy
accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)

# Calculate precision
precision <- confusion_matrix[2, 2] / sum(confusion_matrix[, 2])

# Calculate sensitivity (true positive rate)
sensitivity <- confusion_matrix[2, 2] / sum(confusion_matrix[2, ])

# Calculate specificity (true negative rate)
specificity <- confusion_matrix[1, 1] / sum(confusion_matrix[1, ])

# Print the results
cat("Accuracy:", accuracy, "\n")
cat("Precision:", precision, "\n")
cat("Sensitivity:", sensitivity, "\n")
cat("Specificity:", specificity, "\n")

```

### this time we split the data into training and testing sets using a 80-20 ratio.
### then we created a formula for the decision tree using the specified class attribute. Then, built the decision tree model using the `rpart()` function with the "class" method and information gain as the splitting criterion.


```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (80%) and testing (20%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.8 * nrow(CreditCard))
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using information gain on the training data
tree_info_gain <- rpart(formula, data = train_data, method = "class", parms = list(split = "information"))

# Visualize the decision tree
rpart.plot(tree_info_gain)

# Make predictions on the test data
predicted <- predict(tree_info_gain, test_data, type = "class")

# Calculate accuracy
accuracy <- sum(predicted == test_data$Approved) / nrow(test_data)
cat("Accuracy with Information Gain (80-20 Split)", accuracy, "\n")


```

### we again split the data into training and testing sets using a 80-20 ratio. 
### then, we Built the decision tree using C5.0 with gain ratio Assuming that “Approved" is a binary variable (0 or 1)  and printed it.

```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (80%) and testing (20%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.8 * nrow(CreditCard))
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using gain ratio on the training data
tree_gain_ratio <- rpart(formula, data = train_data, method = "class", parms = list(split = "gain"))

# Visualize the decision tree
rpart.plot(tree_gain_ratio)

# Make predictions on the test data
predicted <- predict(tree_gain_ratio, test_data, type = "class")

# Calculate accuracy
accuracy <- sum(predicted == test_data$Approved) / nrow(test_data)
cat("Accuracy with Gain Ratio (80-20 Split):", accuracy, "\n")


```


### we again split the data  using a 80-20 ratio. 
### then created a formula for the decision tree using the specified class attribute. Then, built the decision tree model using the `rpart()` function with the "class" method and Gini index as the splitting criterion.

```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (80%) and testing (20%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.8 * nrow(CreditCard))
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using Gini index on the training data
tree_gini_index <- rpart(formula, data = train_data, method = "class", control = rpart.control(cp = 0.005))

# Visualize the decision tree
rpart.plot(tree_gini_index)

# Make predictions on the test data
predicted <- predict(tree_gini_index, test_data, type = "class")

# Calculate accuracy
accuracy <- sum(predicted == test_data$Approved) / nrow(test_data)
cat("Accuracy with Gini Index (80-20 Split):", accuracy, "\n")

```
## after doing 3 trees for 80-20 split the best tree with highest accuracy was Gain ratio


### we made predictions on the test data
### created the Confusion Matrix
### calculated the Accuracy, Precision, ensitivity (True Positive Rate) , Specificity (True Negative Rate) and printed the results.


```{r}
tree <- rpart(Approved ~ ., data = train_data, method = "class")
# Make predictions on the test data
predicted <- predict(tree, test_data, type = "class")

# Calculate confusion matrix
confusion_matrix <- table(Actual = test_data$Approved, Predicted = predicted)

# Calculate accuracy
accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)

# Calculate precision
precision <- confusion_matrix[2, 2] / sum(confusion_matrix[, 2])

# Calculate sensitivity (true positive rate)
sensitivity <- confusion_matrix[2, 2] / sum(confusion_matrix[2, ])

# Calculate specificity (true negative rate)
specificity <- confusion_matrix[1, 1] / sum(confusion_matrix[1, ])

# Print the results
cat("Accuracy:", accuracy, "\n")
cat("Precision:", precision, "\n")
cat("Sensitivity:", sensitivity, "\n")
cat("Specificity:", specificity, "\n")

```

### and this time we split the data into training and testing sets using a 60-40 ratio.
### then we created a formula for the decision tree using the specified class attribute. Then, built the decision tree model using the `rpart()` function with the "class" method and information gain as the splitting criterion.

```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (60%) and testing (40%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.6 * nrow(CreditCard))
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using information gain on the training data
tree_info_gain <- rpart(formula, data = train_data, method = "class", parms = list(split = "information"))

# Visualize the decision tree
rpart.plot(tree_info_gain)

# Make predictions on the test data
predicted <- predict(tree_info_gain, test_data, type = "class")

# Calculate accuracy
accuracy <- sum(predicted == test_data$Approved) / nrow(test_data)
cat("Accuracy with Information Gain (60-40 Split):", accuracy, "\n")


```

### we again split the data into training and testing sets using a 60-40 ratio. 
### then, we Built the decision tree using C5.0 with gain ratio Assuming that “Approved" is a binary variable (0 or 1)  and printed it.

```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (60%) and testing (40%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.6 * nrow(CreditCard))
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using gain ratio on the training data
tree_gain_ratio <- rpart(formula, data = train_data, method = "class", parms = list(split = "gain"))

# Visualize the decision tree
rpart.plot(tree_gain_ratio)

# Make predictions on the test data
predicted <- predict(tree_gain_ratio, test_data, type = "class")

# Calculate accuracy
accuracy <- sum(predicted == test_data$Approved) / nrow(test_data)
cat("Accuracy with Gain Ratio (60-40 Split):", accuracy, "\n")

```

### we again again split the data  using a 60-40 ratio. 
### then created a formula for the decision tree using the specified class attribute. Then, built the decision tree model using the `rpart()` function with the "class" method and Gini index as the splitting criterion.

```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (60%) and testing (40%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.6 * nrow(CreditCard))
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using Gini index on the training data
tree_gini_index <- rpart(formula, data = train_data, method = "class", control = rpart.control(cp = 0.005))

# Visualize the decision tree
rpart.plot(tree_gini_index)

# Make predictions on the test data
predicted <- predict(tree_gini_index, test_data, type = "class")

# Calculate accuracy
accuracy <- sum(predicted == test_data$Approved) / nrow(test_data)
cat("Accuracy with Gini Index (60-40 Split):", accuracy, "\n")

```
## after doing 3 trees for 60-40 split the best tree with highest accuracy was information gain and Gain ratio


### here we made predictions on the test data
### created the Confusion Matrix
### calculated the Accuracy, Precision, ensitivity (True Positive Rate) , Specificity (True Negative Rate) and printed the results. 

```{r}
tree <- rpart(Approved ~ ., data = train_data, method = "class")
# Make predictions on the test data
predicted <- predict(tree, test_data, type = "class")

# Calculate confusion matrix
confusion_matrix <- table(Actual = test_data$Approved, Predicted = predicted)

# Calculate accuracy
accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)

# Calculate precision
precision <- confusion_matrix[2, 2] / sum(confusion_matrix[, 2])

# Calculate sensitivity (true positive rate)
sensitivity <- confusion_matrix[2, 2] / sum(confusion_matrix[2, ])

# Calculate specificity (true negative rate)
specificity <- confusion_matrix[1, 1] / sum(confusion_matrix[1, ])

# Print the results
cat("Accuracy:", accuracy, "\n")
cat("Precision:", precision, "\n")
cat("Sensitivity:", sensitivity, "\n")
cat("Specificity:", specificity, "\n")

```
## after calculate the accuracy for each split, the highest accuracy was 80-20 split





### to determine the appropriate number of clusters for our data, we need to observe the elbow point.
### we Selected the Numeric Attributes for Clustering
### then we specified The range of K values to be tested in the elbow method
### we initialized  An empty vector (wcss_values) to store the Within-Cluster Sum of Squares (WCSS) values for each tested K.
### then calculated the WCSS for Each K Value.
### then created a data frame (elbow_data) to store the K values and their corresponding WCSS values.
### finally we Ploted K versus WCSS Using ggplot2


```{r}
# Load required libraries
library(readxl)
library(cluster)
library(ggplot2)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Select the numeric attributes for clustering (exclude non-numeric or irrelevant columns)
numeric_attributes <- CreditCard[, c("Age", "YearsEmployed", "Debt", "CreditScore", "Income")]

# Specify the range of K values you want to test
k_values <- 1:10  # You can adjust the range

# Initialize a vector to store the WCSS values
wcss_values <- vector()

# Calculate WCSS for each K value
for (k in k_values) {
  kmeans_result <- kmeans(numeric_attributes, centers = k)
  wcss_values <- c(wcss_values, kmeans_result$tot.withinss)
}

# Create a data frame to plot K versus WCSS
elbow_data <- data.frame(K = k_values, WCSS = wcss_values)

# Plot K versus WCSS
ggplot(elbow_data, aes(x = K, y = WCSS)) +
  geom_line() +
  geom_point() +
  labs(x = "Number of Clusters (K)", y = "Within-Cluster Sum of Squares (WCSS)") +
  ggtitle("Elbow Method to Choose K")

```




```{r}
# Load required libraries
library(dplyr)
library(factoextra)


# Encode categorical variables (Industry, Ethnicity, Citizen)
CreditCard_encoded <- CreditCard %>%
  mutate(
    Industry = as.numeric(factor(Industry)),
    Ethnicity = as.numeric(factor(Ethnicity)),
    Citizen = as.numeric(factor(Citizen))
  )

# Extract the features for clustering (excluding 'Approved' column)
features <- CreditCard_encoded %>%
  select(-Approved)

# Scale the features
scaled_data <- scale(features)

# Perform k-means clustering with k=2
kmeans_result <- kmeans(scaled_data, centers = 2, nstart = 25)

# Visualize the clusters
fviz_cluster(kmeans_result, data = scaled_data, geom = "point",
             stand = FALSE, # Don't standardize the variables
             main = "K-Means Clustering (k=2)")


```

### the following code will calculate the Bcubed precision and recall, witch will assess the quality of our clustering result by comparing it to known ground truth labels.

```{r}

cluster_assignments <- kmeans_result$cluster
ground_truth_labels <- CreditCard$Approved 
data <- data.frame(cluster = cluster_assignments, label = ground_truth_labels)

# Function to calculate BCubed precision and recall
calculate_bcubed_metrics <- function(data) {
  n <- nrow(data)
  precision_sum <- 0
  recall_sum <- 0

  for (i in 1:n) {
    cluster <- data$cluster[i]
    label <- data$label[i]
    
    # Count the number of items from the same category within the same cluster
    same_category_same_cluster <- sum(data$label[data$cluster == cluster] == label)
    
    # Count the total number of items in the same cluster
    total_same_cluster <- sum(data$cluster == cluster)
    
    # Count the total number of items with the same category
    total_same_category <- sum(data$label == label)
    
    # Calculate precision and recall for the current item and add them to the sums
    precision_sum <- precision_sum + same_category_same_cluster / total_same_cluster
    recall_sum <- recall_sum + same_category_same_cluster / total_same_category
  }

  # Calculate average precision and recall
  precision <- precision_sum / n
  recall <- recall_sum / n

  return(list(precision = precision, recall = recall))
}

# Calculate BCubed precision and recall
metrics <- calculate_bcubed_metrics(data)

# Extract precision and recall from the metrics
precision <- metrics$precision
recall <- metrics$recall

# Print the results
cat("BCubed Precision:", precision, "\n")
cat("BCubed Recall:", recall, "\n")

```

### the following code performs k-means clustering on our dataset after encoding and scaling the features. It also calculates the Silhouette Score and the Total Within-Cluster Sum of Squares (WSS) for evaluating the clustering results.

```{r}
# Load required libraries
library(cluster)
library(factoextra)
library(dplyr)


# Encoding and scaling
CreditCard_encoded <- CreditCard %>%
  mutate(
    Industry = as.numeric(factor(Industry)),
    Ethnicity = as.numeric(factor(Ethnicity)),
    Citizen = as.numeric(factor(Citizen))
  )

# Remove the class label 'Approved'
CreditCard_scaled <- CreditCard_encoded %>%
  select(-Approved) %>%
  scale()

# Specify the number of clusters (k)
k <- 2

# Perform k-means clustering with k clusters
kmeans_result <- kmeans(CreditCard_scaled, centers = k, nstart = 25)

# Silhouette Score
silhouette_score <- silhouette(kmeans_result$cluster, dist(CreditCard_scaled))
cat("Silhouette Score for k =", k, ":", mean(silhouette_score[, "sil_width"]), "\n")

# Total Within-Cluster Sum of Squares (WSS)
wss <- kmeans_result$tot.withinss
cat("Total Within-Cluster Sum of Squares (WSS) for k =", k, ":", wss, "\n")


```


### the following code performs k-means clustering our dataset with a specified number of clusters (k = 2). It then calculates silhouette information, including the average silhouette width for each cluster, and visualizes the silhouette plot using the factoextra library.

```{r}
# Load required libraries
library(cluster)
library(factoextra)
library(dplyr)
library(fpc)  # Add this line to load the fpc package

# Assume 'CreditCard' is your dataset

# Encoding and scaling
CreditCard_encoded <- CreditCard %>%
  mutate(
    Industry = as.numeric(factor(Industry)),
    Ethnicity = as.numeric(factor(Ethnicity)),
    Citizen = as.numeric(factor(Citizen))
  )

# Remove the class label 'Approved'
CreditCard_scaled <- CreditCard_encoded %>%
  select(-Approved) %>%
  scale()

# Specify the number of clusters (k)
k <- 2

# Perform k-means clustering with k clusters
kmeans_result <- kmeans(CreditCard_scaled, centers = k, nstart = 25)

# Calculate silhouette information
sil_info <- silhouette(kmeans_result$cluster, dist(CreditCard_scaled))

# Calculate average silhouette width for each cluster
avg_sil_width <- cluster.stats(dist(CreditCard_scaled), kmeans_result$cluster)$avg.silwidth

# Visualize silhouette plot
fviz_silhouette(sil_info)

# Print average silhouette width for each cluster
cat("Average Silhouette Width for each cluster:\n")
cat(paste("Cluster", 1:k, ": ", avg_sil_width, "\n"))


```




### the following code performs k-means clustering on our dataset with a specified number of clusters (k = 3). It then visualizes the clusters using the factoextra library. 

```{r}
# Load required libraries
library(dplyr)
library(factoextra)


# Encode categorical variables (Industry, Ethnicity, Citizen)
CreditCard_encoded <- CreditCard %>%
  mutate(
    Industry = as.numeric(factor(Industry)),
    Ethnicity = as.numeric(factor(Ethnicity)),
    Citizen = as.numeric(factor(Citizen))
  )

# Extract the features for clustering (excluding 'Approved' column)
features <- CreditCard_encoded %>%
  select(-Approved)

# Scale the features
scaled_data <- scale(features)

# Perform k-means clustering with k=3
kmeans_result <- kmeans(scaled_data, centers = 3, nstart = 25)

# Visualize the clusters
fviz_cluster(kmeans_result, data = scaled_data, geom = "point",
             stand = FALSE, # Don't standardize the variables
             main = "K-Means Clustering (k=3)")

```
### the following code performs k-means clustering on our dataset with a specified number of clusters (k = 3). It then calculates the Silhouette Score and the Total Within-Cluster Sum of Squares (WSS) for evaluating the clustering results.

```{r}
 cluster_assignments <- kmeans_result$cluster
ground_truth_labels <- CreditCard$Approved 
data <- data.frame(cluster = cluster_assignments, label = ground_truth_labels)

# Function to calculate BCubed precision and recall
calculate_bcubed_metrics <- function(data) {
  n <- nrow(data)
  precision_sum <- 0
  recall_sum <- 0

  for (i in 1:n) {
    cluster <- data$cluster[i]
    label <- data$label[i]
    
    # Count the number of items from the same category within the same cluster
    same_category_same_cluster <- sum(data$label[data$cluster == cluster] == label)
    
    # Count the total number of items in the same cluster
    total_same_cluster <- sum(data$cluster == cluster)
    
    # Count the total number of items with the same category
    total_same_category <- sum(data$label == label)
    
    # Calculate precision and recall for the current item and add them to the sums
    precision_sum <- precision_sum + same_category_same_cluster / total_same_cluster
    recall_sum <- recall_sum + same_category_same_cluster / total_same_category
  }

  # Calculate average precision and recall
  precision <- precision_sum / n
  recall <- recall_sum / n

  return(list(precision = precision, recall = recall))
}

# Calculate BCubed precision and recall
metrics <- calculate_bcubed_metrics(data)

# Extract precision and recall from the metrics
precision <- metrics$precision
recall <- metrics$recall

# Print the results
cat("BCubed Precision:", precision, "\n")
cat("BCubed Recall:", recall, "\n")

```
### the following code performs k-means clustering on our dataset with a specified number of clusters (k = 3). It then calculates silhouette information, including the average silhouette width for each cluster, and visualizes the silhouette plot using the factoextra library.

```{r}
# Load required libraries
library(cluster)
library(factoextra)
library(dplyr)

# Encoding and scaling
CreditCard_encoded <- CreditCard %>%
  mutate(
    Industry = as.numeric(factor(Industry)),
    Ethnicity = as.numeric(factor(Ethnicity)),
    Citizen = as.numeric(factor(Citizen))
  )

# Remove the class label 'Approved'
CreditCard_scaled <- CreditCard_encoded %>%
  select(-Approved) %>%
  scale()

# Specify the number of clusters (k)
k <- 3

# Perform k-means clustering with k clusters
kmeans_result <- kmeans(CreditCard_scaled, centers = k, nstart = 25)

# Silhouette Score
silhouette_score <- silhouette(kmeans_result$cluster, dist(CreditCard_scaled))
cat("Silhouette Score for k =", k, ":", mean(silhouette_score[, "sil_width"]), "\n")

# Total Within-Cluster Sum of Squares (WSS)
wss <- kmeans_result$tot.withinss
cat("Total Within-Cluster Sum of Squares (WSS) for k =", k, ":", wss, "\n")



```

### the following code performs k-means clustering on our dataset with a specified number of clusters (k = 3). It then calculates silhouette information, including the average silhouette width for each cluster, and visualizes the silhouette plot using the factoextra library.

```{r}
# Load required libraries
library(cluster)
library(factoextra)
library(dplyr)


# Encoding and scaling
CreditCard_encoded <- CreditCard %>%
  mutate(
    Industry = as.numeric(factor(Industry)),
    Ethnicity = as.numeric(factor(Ethnicity)),
    Citizen = as.numeric(factor(Citizen))
  )

# Remove the class label 'Approved'
CreditCard_scaled <- CreditCard_encoded %>%
  select(-Approved) %>%
  scale()

# Specify the number of clusters (k)
k <- 3

# Perform k-means clustering with k clusters
kmeans_result <- kmeans(CreditCard_scaled, centers = k, nstart = 25)

# Calculate silhouette information
sil_info <- silhouette(kmeans_result$cluster, dist(CreditCard_scaled))

# Calculate average silhouette width for each cluster
avg_sil_width <- cluster.stats(dist(CreditCard_scaled), kmeans_result$cluster)$avg.silwidth

# Visualize silhouette plot
fviz_silhouette(sil_info)

# Print average silhouette width for each cluster
cat("Average Silhouette Width for each cluster:\n")
cat(paste("Cluster", 1:k, ": ", avg_sil_width, "\n"))


```

### code performs k-means clustering on our dataset with a specified number of clusters (k = 4). It then visualizes the clusters using the factoextra library

```{r}
# Load required libraries
library(dplyr)
library(factoextra)


# Encode categorical variables (Industry, Ethnicity, Citizen)
CreditCard_encoded <- CreditCard %>%
  mutate(
    Industry = as.numeric(factor(Industry)),
    Ethnicity = as.numeric(factor(Ethnicity)),
    Citizen = as.numeric(factor(Citizen))
  )

# Extract the features for clustering (excluding 'Approved' column)
features <- CreditCard_encoded %>%
  select(-Approved)

# Scale the features
scaled_data <- scale(features)

# Perform k-means clustering with k=2
kmeans_result <- kmeans(scaled_data, centers = 4, nstart = 25)

# Visualize the clusters
fviz_cluster(kmeans_result, data = scaled_data, geom = "point",
             stand = FALSE, # Don't standardize the variables
             main = "K-Means Clustering (k=4)")

```
## the following code calculates BCubed precision and recall for evaluating the quality of our clustering result compared to ground truth labels.
```{r}
cluster_assignments <- kmeans_result$cluster
ground_truth_labels <- CreditCard$Approved 
data <- data.frame(cluster = cluster_assignments, label = ground_truth_labels)

# Function to calculate BCubed precision and recall
calculate_bcubed_metrics <- function(data) {
  n <- nrow(data)
  precision_sum <- 0
  recall_sum <- 0

  for (i in 1:n) {
    cluster <- data$cluster[i]
    label <- data$label[i]
    
    # Count the number of items from the same category within the same cluster
    same_category_same_cluster <- sum(data$label[data$cluster == cluster] == label)
    
    # Count the total number of items in the same cluster
    total_same_cluster <- sum(data$cluster == cluster)
    
    # Count the total number of items with the same category
    total_same_category <- sum(data$label == label)
    
    # Calculate precision and recall for the current item and add them to the sums
    precision_sum <- precision_sum + same_category_same_cluster / total_same_cluster
    recall_sum <- recall_sum + same_category_same_cluster / total_same_category
  }

  # Calculate average precision and recall
  precision <- precision_sum / n
  recall <- recall_sum / n

  return(list(precision = precision, recall = recall))
}

# Calculate BCubed precision and recall
metrics <- calculate_bcubed_metrics(data)

# Extract precision and recall from the metrics
precision <- metrics$precision
recall <- metrics$recall

# Print the results
cat("BCubed Precision:", precision, "\n")
cat("BCubed Recall:", recall, "\n")
```

### the following performs k-means clustering on our dataset with a specified number of clusters (k = 4). It then calculates the Silhouette Score and the Total Within-Cluster Sum of Squares (WSS) for evaluating the clustering results

```{r}
# Load required libraries
library(cluster)
library(factoextra)
library(dplyr)


# Encoding and scaling
CreditCard_encoded <- CreditCard %>%
  mutate(
    Industry = as.numeric(factor(Industry)),
    Ethnicity = as.numeric(factor(Ethnicity)),
    Citizen = as.numeric(factor(Citizen))
  )

# Remove the class label 'Approved'
CreditCard_scaled <- CreditCard_encoded %>%
  select(-Approved) %>%
  scale()

# Specify the number of clusters (k)
k <- 4

# Perform k-means clustering with k clusters
kmeans_result <- kmeans(CreditCard_scaled, centers = k, nstart = 25)

# Silhouette Score
silhouette_score <- silhouette(kmeans_result$cluster, dist(CreditCard_scaled))
cat("Silhouette Score for k =", k, ":", mean(silhouette_score[, "sil_width"]), "\n")


# Total Within-Cluster Sum of Squares (WSS)
wss <- kmeans_result$tot.withinss
cat("Total Within-Cluster Sum of Squares (WSS) for k =", k, ":", wss, "\n")


```

### the following code performs k-means clustering on our dataset with a specified number of clusters (k = 4). It then calculates the average silhouette width for each cluster and visualizes the silhouette plot using the factoextra library.

```{r}
# Load required libraries
library(cluster)
library(factoextra)
library(dplyr)


# Encoding and scaling
CreditCard_encoded <- CreditCard %>%
  mutate(
    Industry = as.numeric(factor(Industry)),
    Ethnicity = as.numeric(factor(Ethnicity)),
    Citizen = as.numeric(factor(Citizen))
  )

# Remove the class label 'Approved'
CreditCard_scaled <- CreditCard_encoded %>%
  select(-Approved) %>%
  scale()

# Specify the number of clusters (k)
k <- 4

# Perform k-means clustering with k clusters
kmeans_result <- kmeans(CreditCard_scaled, centers = k, nstart = 25)


# Calculate average silhouette width for each cluster
avg_sil_width <- cluster.stats(dist(CreditCard_scaled), kmeans_result$cluster)$avg.silwidth

# Visualize silhouette plot
fviz_silhouette(sil_info)

# Print average silhouette width for each cluster
cat("Average Silhouette Width for each cluster:\n")
cat(paste("Cluster", 1:k, ": ", avg_sil_width, "\n"))


```



### the following code performs the silhouette method to determine the optimal number of clusters for k-means clustering. It uses a range of cluster numbers (k_range) and calculates the average silhouette width for each.

```{r}
# Load required libraries
library(cluster)
library(factoextra)
library(dplyr)

# Encoding and scaling
CreditCard_encoded <- CreditCard %>%
  mutate(
    Industry = as.numeric(factor(Industry)),
    Ethnicity = as.numeric(factor(Ethnicity)),
    Citizen = as.numeric(factor(Citizen))
  )

# Remove the class label 'Approved'
CreditCard_scaled <- CreditCard_encoded %>%
  select(-Approved) %>%
  scale()

# Specify the range of clusters you want to explore
k_range <- 2:10

# Function to calculate average silhouette width for a given number of clusters (k)
calculate_silhouette <- function(k) {
  kmeans_result <- kmeans(CreditCard_scaled, centers = k, nstart = 25)
  silhouette_score <- silhouette(kmeans_result$cluster, dist(CreditCard_scaled))
  return(mean(silhouette_score[, "sil_width"]))
}

# Calculate silhouette width for each k and store in a vector
sil_widths <- sapply(k_range, calculate_silhouette)

# Find the optimal number of clusters based on the maximum silhouette width
optimal_k_sil <- k_range[which.max(sil_widths)]

# Compare results
cat("Optimal number of clusters (silhouette method):", optimal_k_sil, "\n")

# Plot silhouette width
plot(k_range, sil_widths, type = "b", pch = 19, frame = FALSE, 
     xlab = "Number of Clusters (k)", ylab = "Average Silhouette Width",
     main = "Silhouette Method for Optimal Number of Clusters")

# Add a point for the optimal k
points(optimal_k_sil, sil_widths[optimal_k_sil - min(k_range) + 1], col = "red", cex = 2, pch = 19)


```
## after calculate the average silhouette width the best cluster was k=2



