---
title: "R Notebook"
output: html_notebook
---
This is our dataset before pre-processing.
```{r}
library(readxl)
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")
print(CreditCard)
```
Data cleaning: our data set already cleaned, there are no incomplete all colomns are filled , inconsistent each colomn have the same type  or intentional data. it has only noisy data (outliers) and we removed them.






Scatter plot shows the relationship between the age and the yearsEmployed if the age increased the years also will increase.
```{r}
#Scatter plot
with(CreditCard, plot(Age,YearsEmployed))
```

Hisogram shows the most frequent debt which is -1.
```{r}
#histogram 
var(CreditCard$Debt)
hist(CreditCard$Debt)
```

Barplot shows how many applicants are citizen or not.
```{r}
#barplot
library(dplyr)
CreditCard$Citizen %>% table() %>% barplot()
```
*Data intgration(correlation analysis) first we will compare the nominal colomns with the class label.

Data intgration(correlation analysis) between the Industry colomn (nominal data type) and the class label colomn approved. 
We calculated the chi square = 98.323 and the df= 13 so we found that the crtical value for df 13 is = 34.528 ,the chi square is larger than the critical value we reject H0 so they are depandent and correlated .
```{r}
contingency_table <- table(CreditCard$Industry, CreditCard$Approved)
chi_square_result <- chisq.test(contingency_table)
print(chi_square_result)
``` 
Data intgration(correlation analysis) between the Ethnicity colomn (nominal data type) and the class label colomn approved. 
We calculated the chi square = 41.813 and the df= 4 so we found that the crtical value for df 4 is = 18.465 ,the chi square is larger than the critical value we reject H0 so they are depandent and correlated .
```{r}
contingency_table <- table(CreditCard$Ethnicity, CreditCard$Approved)
chi_square_result <- chisq.test(contingency_table)
print(chi_square_result)
```
Data intgration(correlation analysis) between the Citizen colomn (nominal data type) and the class label colomn approved. 
We calculated the chi square = 9.1916 and the df= 2 so we found that the crtical value for df 2 is = 13.815 ,the chi square is larger than the critical value we reject H0 so they are depandent and correlated .

```{r}
contingency_table <- table(CreditCard$Citizen, CreditCard$Approved)
chi_square_result <- chisq.test(contingency_table)
print(chi_square_result)
```
after we did chi-square all the columns was highly correlated to we didn't apply feature selection.






*Data intgration(correlation analysis) Second we will compare the numeric colomns with the class label,we found that all numeric coulmns are positive corrleated except zip code coulmn so we need to deleate it.
```{r}
correlation<- cor(CreditCard$Age, CreditCard$Approved)
print(correlation)
```

```{r}
correlation<- cor(CreditCard$Debt, CreditCard$Approved)
print(correlation)
```

```{r}
correlation<- cor(CreditCard$YearsEmployed, CreditCard$Approved)
print(correlation)
```

```{r}
correlation<- cor(CreditCard$CreditScore, CreditCard$Approved)
print(correlation)
```
```{r}
correlation<- cor(CreditCard$Income, CreditCard$Approved)
print(correlation)
```


```{r}
correlation<- cor(CreditCard$ZipCode, CreditCard$Approved)
print(correlation)
```
```{r}
CreditCard<-subset(CreditCard,select= -ZipCode)
print(CreditCard)
```



Data cleaning (noisy outliers)
We relize in Debt coulmn has wide range so it might have outliers
First we detect and found the outliers from the debt colomn .
```{r}
#Find outliers

library(outliers)

OutDebt = outlier(CreditCard$Debt, logical =TRUE)
sum(OutDebt)
Find_outlier = which(OutDebt ==TRUE, arr.ind = TRUE)
OutDebt
Find_outlier
```
Second we deleted the outliers, and then printed the data set after the changes.
```{r}
#remove outliers

CreditCard= CreditCard[-Find_outlier,]
print(CreditCard)
```
Data cleaning (noisy outliers)
As same as Debt ,Income has also huge range.
Here also we detect and found the outliers from the income colomn. .
```{r}
#Find outliers for Income

library(outliers)

OutIncome = outlier(CreditCard$Income, logical =TRUE)
sum(OutIncome)
Find_outlier = which(OutIncome ==TRUE, arr.ind = TRUE)
OutIncome
Find_outlier
```
Second we deleted the outliers from the income colomn, and then printed the data set after the changes.

```{r}
#remove outliers

CreditCard= CreditCard[-Find_outlier,]
print(CreditCard)
```
Data transformation (normalization)
The range of Debt was very wide so we did normalization on Debt to make it in smaller range.
```{r}
#normlization
CreditCard [, 3:3] = scale(CreditCard [, 3:3])
print(CreditCard)

```
Data transformation (encoding)
we noticed that the Income has huge range and The mean = 1017.386 so any value bigger than the mean we encoded it to high and any value smaller than the mean it became low.
```{r}
#Encoding for income
Income_mean<- mean(CreditCard$Income)
print(Income_mean)
CreditCard$Income[CreditCard$Income > Income_mean]<- "High"
CreditCard$Income[CreditCard$Income <= Income_mean]<- "low"
print(CreditCard)
```

This is our Dataset after pre-processing.
```{r}
print(CreditCard)
```
```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (70%) and testing (30%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.7 * nrow(CreditCard))
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using information gain on the training data
tree <- rpart(formula, data = train_data, method = "class", parms = list(split = "information"))

# Visualize the decision tree
rpart.plot(tree)

```




```{r}
# Load required libraries
library(readxl)
library(C50)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (70%) and testing (30%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.7 * nrow(CreditCard))
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Build the decision tree using C5.0 with gain ratio
# Assuming "Approved" is a binary variable (0 or 1)
train_data$Approved <- as.factor(train_data$Approved)
tree <- C5.0(train_data[, -which(names(train_data) == class_attribute)], train_data[[class_attribute]], control = C5.0Control(earlyStopping = FALSE))

# Print the decision tree
plot(tree)

```



```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (70%) and testing (30%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.7 * nrow(CreditCard))
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using Gini index on the training data
tree <- rpart(formula, data = train_data, method = "class", control = rpart.control(cp = 0.005))

# Visualize the decision tree
rpart.plot(tree)


```


```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (80%) and testing (20%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.8 * nrow(CreditCard))  # Use 0.8 for 80% training
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]


# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using information gain on the training data
tree <- rpart(formula, data = train_data, method = "class", parms = list(split = "information"))

# Visualize the decision tree
rpart.plot(tree)

```

```{r}
# Load required libraries
library(readxl)
library(C50)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (80%) and testing (20%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.8 * nrow(CreditCard))  # Use 0.8 for 80% training
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Build the decision tree using C5.0 with gain ratio
# Assuming "Approved" is a binary variable (0 or 1)
train_data$Approved <- as.factor(train_data$Approved)
tree <- C5.0(train_data[, -which(names(train_data) == class_attribute)], train_data[[class_attribute]], control = C5.0Control(earlyStopping = FALSE))

# Print the decision tree
plot(tree)

```


```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (80%) and testing (20%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.8 * nrow(CreditCard))  # Use 0.8 for 80% training
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using Gini index on the training data
tree <- rpart(formula, data = train_data, method = "class", control = rpart.control(cp = 0.005))

# Visualize the decision tree
rpart.plot(tree)

```

```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (90%) and testing (10%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.9 * nrow(CreditCard))  # Use 0.9 for 90% training
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using information gain on the training data
tree <- rpart(formula, data = train_data, method = "class", parms = list(split = "information"))

# Visualize the decision tree
rpart.plot(tree)

```

```{r}
# Load required libraries
library(readxl)
library(C50)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (90%) and testing (10%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.9 * nrow(CreditCard))  # Use 0.9 for 90% training
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Build the decision tree using C5.0 with gain ratio
# Assuming "Approved" is a binary variable (0 or 1)
train_data$Approved <- as.factor(train_data$Approved)
tree <- C5.0(train_data[, -which(names(train_data) == class_attribute)], train_data[[class_attribute]], control = C5.0Control(earlyStopping = FALSE))

# Print the decision tree
plot(tree)

```

```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(rpart)
library(rpart.plot)

# Read the dataset
CreditCard <- read_excel("C:/Users/Noni/Desktop/Datamining project/Credit Card Approvals/Dataset/CreditCard.xlsx")

# Specify the class attribute
class_attribute <- "Approved"

# List of character attributes to encode
char_attributes_to_encode <- c("Industry", "Ethnicity", "Citizen")

# Label Encoding
for (char_attribute in char_attributes_to_encode) {
  CreditCard[[char_attribute]] <- as.numeric(factor(CreditCard[[char_attribute]]))
}

# Split the data into training (90%) and testing (10%)
set.seed(123)  # Set seed for reproducibility
sample_index <- sample(1:nrow(CreditCard), 0.9 * nrow(CreditCard))  # Use 0.9 for 90% training
train_data <- CreditCard[sample_index, ]
test_data <- CreditCard[-sample_index, ]

# Create a formula for the decision tree
formula <- as.formula(paste(class_attribute, " ~ ."))

# Build the decision tree using Gini index on the training data
tree <- rpart(formula, data = train_data, method = "class", control = rpart.control(cp = 0.005))

# Visualize the decision tree
rpart.plot(tree)

```


